\section{Background}
\label{sec:bg}

\subsection{Process Models and Their Silences}

Digital forensic process models describe how an investigation should
proceed, not what classes of evidence exist. The DFRWS framework first
codified the seven-phase pipeline of identification, preservation,
collection, examination, analysis, presentation, and decision~\cite{palmer2001roadmap}.
Carrier and Spafford generalized this into a five-phase \emph{Integrated
Digital Investigation Process} that explicitly aligns the digital phases
with the physical-crime-scene phases that
preceded them~\cite{carrier2003getting}. Beebe and Clark extended the
emphasis on hierarchical objectives, arguing that procedural decomposition
must respond to investigative
objectives rather than artifact taxonomy~\cite{beebe2005hierarchical}.
Carrier's later dissertation~\cite{carrier2006hypothesis} formalized
hypothesis testing as the epistemic core of digital investigation: every
analytical claim is a hypothesis that must be falsifiable from the data.

These models are silent on the question we ask. They describe \emph{when}
evidence is acquired, \emph{who} performs the acquisition, and \emph{how} the
chain of custody is preserved. They do not describe \emph{what kinds of
addressing} a piece of evidence supports, or what the architectural
consequences are. The hypothesis-based framework, in particular, presupposes
that an investigator already knows where to look for evidence that might
falsify a hypothesis. The taxonomy we develop is a precondition for that
presupposition: an investigator cannot know where to look until the field
has named what kinds of addressing are available.

\subsection{The Three-Source Folk Model and Its Failure Modes}

The implicit, never-formalized folk model of forensic evidence partitions
sources into three: \emph{disk}, \emph{memory}, and
\emph{log}~\cite{kent2006sp80086,casey2011digital,carrier2005file}. Standard
practitioner training and toolchains -- The Sleuth Kit and
Autopsy~\cite{carrier2005file}, Volatility for memory analysis~\cite{ligh2014artmemory,case2017memory},
and timeline tools such as Plaso/log2timeline~\cite{plaso} -- align with this
partition. The folk model has served the field well for two decades but is
breaking under the weight of two developments.

The first is the rise of \emph{enterprise live response} and endpoint
detection-and-response telemetry. When a Velociraptor agent~\cite{velociraptor}
runs a VQL query and returns a result set of currently-open TCP connections,
that result is not on a disk image; it is not yet in any log; it lives,
briefly, only in the network packet that delivers it from the endpoint to
the investigator. Cohen, Bilby, and Garfinkel argued in
2011~\cite{cohen2011distributed} that this kind of capture is a fundamentally
different forensic operation, but the field continued to treat the resulting
data as ``just another log'' -- an analytical convenience that obscures the
property that the data did not exist before the query was issued and, once
captured and transmitted, cannot be modified by an adversary regardless of
subsequent on-host actions.

The second development is the rise of \emph{content-addressed} ecosystems.
Git~\cite{chacon2014progit}, Sigstore's transparency
log~\cite{newman2022sigstore}, the in-toto framework~\cite{torresarias2019intoto},
the Open Container Initiative image
specification~\cite{ociimagespec}, and IPFS~\cite{benet2014ipfs} all share an
addressing model in which an artifact's identity \emph{is} the cryptographic
hash of its content, and references between artifacts are themselves hashes.
The resulting Merkle DAG has a navigation discipline that maps cleanly onto
none of disk, memory, or log: there is no path, no virtual address, no
timestamp ordering. There is only the hash and the references that hashes
encode. Software supply-chain attacks -- SolarWinds~\cite{cisa2021solarwinds},
the \texttt{event-stream} npm compromise~\cite{ohm2020backstabber}, the
\texttt{xz-utils} backdoor~\cite{cve2024xz} -- forced the forensic community
to reckon with the fact that part of the evidence trail lives in
content-addressed stores that no traditional triage tool understands.

\subsection{Live Response and Endpoint Telemetry}

The architecture of live-response systems makes the address-space question
unavoidable. Velociraptor compiles a VQL expression at the
investigator's console, ships the compiled query to the endpoint, executes
it in the agent process, and streams typed result rows
back~\cite{velociraptor}. The data observed -- which is to say, the rows --
is determined by the query, not by the contents of any file. WMI/WQL on
Windows~\cite{microsoft2023wmi} and OSQuery on
Linux/macOS/Windows~\cite{osquery} share this discipline: the address space
of the evidence is the space of well-formed queries crossed with cursor
tokens. Treating these systems as ``just another log producer'' loses the
key property that the evidence exists \emph{only} because the query was run,
and that the act of running a query at time $t_0$ produces a record that no
adversary acting at $t_1 > t_0$ on the host can alter.

\subsection{Supply Chain Forensics and Content Addressing}

The supply-chain context introduces a different kind of evidence with a
different kind of guarantee. A package built by a CI/CD pipeline has a
content hash; Sigstore's transparency log records the signing event with the
hash and the signing identity; in-toto links record the build steps each
with their input and output hashes~\cite{torresarias2019intoto,newman2022sigstore}.
Investigating a suspicious binary discovered on a compromised endpoint thus
becomes, in part, an exercise in following hashes -- from the binary on disk,
to the OCI layer hash, to the Sigstore entry, to the in-toto link, to the
git commit. Every step is the same operation: hash $\to$ blob $\to$
content graph. This is a navigation primitive, and it has properties
(notably immutability by construction) that the other primitives lack.

\subsection{The Gap}

The literature provides three separate strands -- traditional disk and log
forensics, memory forensics as it matured through Volatility and
Rekall~\cite{ligh2014artmemory,case2017memory}, and emerging work in live
response and supply-chain forensics -- that have not been unified by a
common formal model.

The closest representation-level effort is the CASE/UCO
ontology~\cite{caseuco2018,casey2017advancing}, which provides a typed
vocabulary for describing digital evidence \emph{after} it has been parsed.
Our taxonomy is complementary, not competing: CASE/UCO is \emph{annotation}
(the schema in which findings are exchanged), while the navigation-primitive
taxonomy is \emph{architecture} (the addressing discipline by which findings
are produced). A CASE/UCO record consumed by a downstream tool conveys the
\emph{what}; a navigation-primitive tag (\eg{}, \texttt{nav-primitive: Q})
conveys the \emph{how}, which CASE/UCO does not encode and which determines
the trust profile and durability claims that the consumer can make.

The closest infrastructure-level effort is Carrier's data-abstraction model
and its concrete instantiation in The Sleuth Kit (TSK)~\cite{carrier2005file,tsk_repo}.
TSK exposes \texttt{img\_*}, \texttt{vol\_*}, \texttt{fs\_*}, and \texttt{file\_*}
APIs corresponding to a layered traversal: image $\to$ volume $\to$
filesystem $\to$ file. This layering is foundational for \src{P} navigation
specifically: TSK's \texttt{img\_} layer is what we generalize as CONTAINER,
and TSK's \texttt{fs\_}/\texttt{file\_} layers together constitute the
\src{P}-specialization of NAVIGATE in our framework. The crucial difference
is that TSK's layering is \emph{specific} to disk evidence -- it has no
analog of \texttt{img\_} for memory dumps, no analog of \texttt{vol\_} for
log streams, no analog of \texttt{file\_} for hash-addressed blobs. Our
taxonomy generalizes Carrier's layered insight to four other addressing
schemes that did not exist in their current form when TSK was designed.

Garfinkel's 2009 work on automating disk forensic processing with TSK, XML,
and Python~\cite{garfinkel2009automating} provides early evidence that even
within \src{P} alone the absence of a uniform addressing model forces
ad-hoc adapters: Garfinkel's DFXML output exists precisely because tool-output
schemas were not interoperable. The same problem at five-source scale
motivates the present paper.

Garfinkel's 2010 agenda~\cite{garfinkel2010digital} identified cross-source
correlation as one of the key challenges of the coming decade; sixteen
years later, the field has rich tooling for each source individually but no
formal account of what ``cross-source'' means. We aim to provide that
account.
