\section{Related Work}
\label{sec:related}

We position the navigation-primitive taxonomy against five strands of prior
work: digital-forensic process models, data-abstraction layer models,
representation ontologies, source-specific tooling, and emerging
supply-chain-forensic foundations. In each case our claim is one of
extension or formalization, not displacement.

\subsection{Process Models}

Carrier and Spafford~\cite{carrier2003getting} establish the Integrated
Digital Investigation Process and, in the same paper, argue for the value
of a \emph{physical} grounding: digital evidence has a physical substrate
(a disk, a memory chip), and the investigation should respect that. The
phases they articulate -- readiness, deployment, physical scene, digital
scene, review -- describe the temporal arc of an investigation. Our work
is orthogonal: we do not displace the phases; we observe that within the
\emph{digital scene} phase, evidence has not just one underlying
representation (their bytes-on-a-disk model) but at least five distinct
addressing schemes, each of which has independent epistemological
properties. Where Carrier and Spafford's model says \emph{when} digital
evidence is collected, ours says \emph{how} it is reached.

Carrier's later hypothesis-based framework~\cite{carrier2006hypothesis}
provides the epistemic engine: every analytical claim is a hypothesis with
a class of falsifying observations. Our taxonomy supplies a partial answer
to a question the hypothesis framework leaves open: \emph{where, in
principle, can falsifying observations live?} They live in the address
spaces of the five primitives. The Q-Durability and C-Intrinsic-Durability
properties (Section~\ref{sec:properties}) restrict the kinds of
falsifications that an adversary can stage post-hoc.

Beebe and Clark~\cite{beebe2005hierarchical} contribute the hierarchical
objectives-based framework, which modulates the procedural decomposition by
investigative goal. Our work is again complementary: a framework that
selects analytical objectives benefits from a taxonomy that names the
addressing modalities the objectives may need to reach. The DFRWS roadmap
paper~\cite{palmer2001roadmap} and the NIST Special Publication
800-86~\cite{kent2006sp80086} similarly provide procedural guidance that
our taxonomy informs but does not replace.

\subsection{Data-Abstraction Layer Models}

Carrier's data-abstraction model~\cite{carrier2005file} layers digital
evidence into physical media, volume, file system, file content, and
application abstraction levels. The model is foundational for understanding
\emph{what abstraction level} a particular artifact inhabits. Our taxonomy
is orthogonal: it asks \emph{which addressing mechanism} is required to
reach the bytes at any given level. Carrier's vertical axis (which
abstraction is being viewed) and our horizontal axis (which navigation
primitive is in play) jointly span a 2-D classification space whose cells
correspond to concrete forensic operations. For example, the cell
(\src{P}, file-system level) is what \texttt{ext4fs-forensic} operates in;
(\src{M}, application level) is what a memory carver targeting Chrome's
heap operates in; (\src{C}, application level) is what
\texttt{sigstore-forensic} targets.

McKemmish's foundational definition of forensic computing~\cite{mckemmish1999forensic}
identifies preservation, identification, extraction, documentation, and
interpretation as the activities of the discipline. Forensic computing as
characterized in 1999 addressed primarily \src{P} and \src{L}. We extend
the discipline by formally admitting \src{M}, \src{Q}, and \src{C} as
first-class source types -- not as variant flavors of disk evidence, but as
sources whose addressing schemes and durability properties differ
fundamentally.

\subsection{Representation Ontologies}

The Cyber-investigation Analysis Standard Expression (CASE), with its UCO
(Unified Cyber-investigation Ontology) foundation~\cite{caseuco2018,casey2017advancing},
provides a vocabulary for representing digital evidence in a sharable, tool-agnostic form.
CASE/UCO operates at the \emph{representation} layer:
once an artifact has been identified, parsed, and described, CASE provides
the linguistic structure to communicate it. Our taxonomy operates at the
\emph{infrastructure} layer: it is concerned with the architecture by which
artifacts are reached and parsed in the first place. The two are
complementary. A CASE/UCO record can carry an additional \emph{source-type}
attribute populated from our taxonomy, which provides downstream consumers
of CASE data with the trust-relevant provenance information that the
representation alone does not encode.

\subsection{Source-Specific Tooling and Frameworks}

The Sleuth Kit and Autopsy~\cite{carrier2005file} are the canonical
\src{P}-source toolkits, exposing a path-based filesystem navigation API
over disk image containers. Volatility~\cite{ligh2014artmemory} and
Volatility~3~\cite{volatility3}, together with Rekall~\cite{rekall2016},
are the canonical \src{M}-source toolkits, with EPROCESS-walking and
page-table-resolving APIs. Plaso/log2timeline~\cite{plaso} is the canonical
\src{L}-source toolkit, normalizing dozens of log formats into a unified
timeline schema. The work of Case and Richard~\cite{case2017memory}
chronicled the maturation of memory forensics over the
preceding decade. Garfinkel's earlier work on file-carving formalized
recovery of artifacts whose filesystem metadata is missing or
unreliable~\cite{garfinkel2007carving}.

These tools and frameworks each excel within a single primitive. Our
taxonomy provides the formal foundation for composing them: by
recognizing that they all produce $\Bstar$ for parser consumption, an
orchestrator can route artifacts to a parser-of-record without committing
to a single tool's record schema. Issen's PARSER crates, for instance,
accept input from a TSK-style mounted volume, a Volatility-style memory
artifact extraction, or a Plaso-style log decode; the dispatch logic does
not care which of those produced the bytes.

\subsection{Live Response and Endpoint Telemetry}

Cohen, Bilby, and Garfinkel~\cite{cohen2011distributed} described GRR Rapid
Response, an enterprise live-response platform that influenced
Velociraptor's later design~\cite{velociraptor}. Their argument was
practical: at enterprise scale, full disk imaging of every potentially
compromised host is infeasible, and live-query collection is the
realistic alternative. They did not, however, propose a formal model of
what live-query evidence \emph{is} or how it differs in kind from disk
evidence. Our \src{Q} primitive and Proposition~\ref{prop:q-durability}
formalize exactly that: live-query evidence is produced by query
execution, not retrieved from storage; once captured and transmitted, it
becomes attacker-durable in a sense disk evidence is not. The OSQuery
project~\cite{osquery} extends the live-query paradigm to a cross-platform
SQL surface; our taxonomy applies uniformly.

\subsection{Supply-Chain and Content-Addressed Foundations}

Torres-Arias \etal{}~\cite{torresarias2019intoto} introduce in-toto, a
framework for end-to-end software supply-chain integrity that records each
build step's input and output by hash. Newman \etal{}'s Sigstore
work~\cite{newman2022sigstore} contributes a public transparency log
(Rekor) for signing events, again addressed by hash. The Open Container
Initiative's image specification~\cite{ociimagespec} establishes
content-addressed image storage. IPFS~\cite{benet2014ipfs} generalizes
content addressing to a peer-to-peer storage substrate.

These projects established content-addressing as a critical primitive for
software-integrity engineering. Our contribution is to incorporate
content-addressing into the forensic source taxonomy as a peer of \src{P},
\src{M}, \src{L}, and \src{Q}, with explicit forensic semantics
(Definition~\ref{def:hash-pivot}) and a durability characterization
(Proposition~\ref{prop:c-durability}). The forensic literature, as far as
we are aware, has not previously elevated \src{C} to a first-class
source type.

\subsection{Closest Prior Work}

The closest prior work is, perhaps surprisingly, Garfinkel's 2010
agenda paper~\cite{garfinkel2010digital}. Garfinkel
identifies cross-source correlation as a foundational unsolved problem and
calls for ``a model that allows multiple sources of evidence to be combined
in principled ways.'' Sixteen years on, the problem remains substantially
unsolved at the formal level. Our taxonomy supplies the missing piece:
classifying sources by navigation primitive yields the formally identified
provenance attribute that a correlation engine needs in order to reason
about trust and to apply the source-appropriate durability profile.
