\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{url}
\usepackage{hyperref}
\usepackage{microtype}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\kern-.125emE\kern-.125emX}}

% --- Theorem-like environments -----------------------------------------------
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}

% --- Macros ------------------------------------------------------------------
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\etal}{\emph{et~al.}}
\newcommand{\etc}{\emph{etc.}}
\newcommand{\src}[1]{\texttt{[#1]}}
\newcommand{\Bstar}{\ensuremath{B^{*}}}

\begin{document}

\title{A Navigation-Primitive Taxonomy of Forensic Evidence Sources:\\
Five Address Spaces and the Architecture of\\
Cross-Source Digital Investigation}

\author{\IEEEauthorblockN{Albert K.T. Hui}
\IEEEauthorblockA{\textit{Security Ronin} \\
albert@securityronin.com}}

\maketitle

\begin{abstract}
Digital forensic frameworks classify evidence by \emph{artifact type} -- what
the evidence \emph{is} -- not by the mechanism required to address and
traverse it. This distinction is consequential: artifact-type classification
does not determine layer dependencies, parser reusability, or
epistemological properties such as attacker-durability. We propose to
classify evidence sources by \emph{navigation primitive} -- the abstract
addressing scheme by which the bytes of evidence are reached -- and show
that exactly five primitives are sufficient to cover contemporary forensic
practice: persistent storage \src{P}, memory \src{M}, log \src{L}, live
query \src{Q}, and content-addressed \src{C}. We formalize evidence sources
as triples $(A, f_{\text{nav}}, f_{\text{parse}})$ where $f_{\text{nav}}: A
\rightharpoonup B^{*}$ is a navigation function valued in byte sequences and
$f_{\text{parse}}: B^{*} \to R$ is a parser valued in records. From this
definition we prove a \emph{Parser Independence} theorem: any two evidence
sources with the same parser produce identical records on identical byte
inputs, regardless of the navigation primitive that retrieved them; we
further establish that the dependency rule (PARSER must not import NAVIGATE)
is the contrapositive of this theorem and is thus a formal architectural
constraint. We define \emph{attacker-durability} of captured evidence and
prove that \src{Q} achieves durability by transmission to investigator
storage (Proposition~\ref{prop:q-durability}) and that \src{C} achieves
durability \emph{intrinsically} from collision-resistance of the addressing
hash (Proposition~\ref{prop:c-durability}); we prove as a corollary that
\src{C} is the unique source type whose durability requires no external
trust anchor. We give a schema-conscious definition of the
\emph{hash-pivot} operation that yields a typed-multiset join across CAS
stores. The taxonomy is hash-agnostic: SHA-1, SHA-256, and future
hash-function migrations are accommodated by a single collision-resistance
parameter. We describe \emph{Issen}, a Rust reference implementation
organized as thirteen crates across five layers (KNOWLEDGE, CONTAINER,
NAVIGATE, PARSER, ORCHESTRATION) whose dependency rules are mechanically
forced by the taxonomy. Four case studies evaluate the framework: a
rootkit-concealed cryptominer requiring \src{P}+\src{M}+\src{Q} fusion; a
supply-chain compromise resolved by hash-pivot through \src{C} stores; an
empirical Parser Independence demonstration with byte-identical
SHA-256 hashes of parser output across four navigation paths into the same
Chrome \texttt{History} file; and a counterfactual case showing concretely
how a \src{P}+\src{M}+\src{L}-only triage tool misses an
\texttt{ld.so.preload} race that \src{Q} captures. The taxonomy supplies
the missing formal foundation for cross-source correlation called for in
Garfinkel's 2010 agenda.
\end{abstract}

\begin{IEEEkeywords}
digital forensics, evidence taxonomy, memory forensics, content-addressed
storage, supply chain forensics, live response, navigation primitive,
parser independence
\end{IEEEkeywords}

\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-taxonomy}
\input{sections/04-properties}
\input{sections/05-implementation}
\input{sections/06-case-studies}
\input{sections/07-related}
\input{sections/08-discussion}
\input{sections/09-conclusion}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
